### 2025-05-05 paper/main.tex
revise the whole text

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{authblk}

\title{Three Algorithms for Merging Hierarchical Navigable Small World Graphs}
\author[1]{Alexander Ponomarenko}
\date{}

% \affil[2]{YugaByte LLC}
\affil[1]{HSE Univerity}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{nameref}
\begin{document}

\maketitle

% В итнтро добавть о том, что такое навигационные графы. Что такое hnsw. Какие операции он поддерживает. 
% Раскрыть глобальную идею мерджа, что merge можно понимать как итеративный процесс из двух уопераций;
% 1. Выбор вершины (мн) для которой мы намериваемся строить окрестность и для какого графа.
% 2. Выбор кандидатов из каждого графа
% 3. Построение окрестности (согласно одной из выбранных стратегий).
% 4. Принятие решение о том, какую информацию передать на следущую итерацию

% TODO для экспериментов. 
% 2. Провести эксперименты для разных размерностей и размеров
% 3. Следить за тем сколько операций "полного поиска было сделано" и сколько операций "local search".
% 5. Сделать все теже эксперементы, но используя в качестве NeighborhoodConstruction k-nn

\section{Introduction}

%TODO здесь надо сказать, именно проважность задачи k-nn для прирожений. Упомянть про RAG, а далее уже перейти к формальной постановке задаче и какие методы для её решения бывают. Далее надо сказать про после того, как было упомянуто про графовые методы, сказать что там етсь задача объединения графов построенных на разных данных. В частности это может быть важно при compaction. 
Navigable small-world graphs have emerged as a cornerstone for efficient approximate nearest neighbor (ANN) search in high-dimensional spaces. This problem is crucial in fields such as recommendation systems, image retrieval, and natural language processing. Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide robust and scalable frameworks for organizing data into graph structures that support fast similarity searches.

\subsection{Nearest Neighbor Search Problem}

% Formally, the nearest neighbor search problem can be defined as follows: Given a set of points $P = \{p_1, p_2, ..., p_n\}$ in a metric space $(X, \rho)$, where $\rho$ is a distance function, and a query point $q \in X$, find the point $p^* \in P$ such that:

% $$p^* = \underset{p \in P}{\mathrm{argmin}} \rho(q, p)$$

Formally, the \(k\)-nearest neighbor search problem can be defined as follows:
Given a set of points
\[
P = \{p_1, p_2, \dots, p_n\}
\]
in a metric space \(\bigl(X, \rho\bigr)\), where \(\rho\) is a distance function, and a query point \(q \in X\), find a set

\[
P_k 
\;=\;
\underset{S \subseteq P,\;|S|=k}{\arg\min}
\;\sum_{p \in S} \rho(q, p)
.
\]
In other words, we seek the set of $k$ points whose total distance to $q$ is minimal.

In high-dimensional spaces, this problem becomes computationally challenging due to the ``curse of dimensionality.'' Exact solutions often degrade to linear scan as dimensionality increases, making them impractical for large-scale applications. This has led to the development of approximate $k$-nearest neighbor (ANN) search methods, which trade perfect accuracy for substantial improvements in computational efficiency.

Three main approaches have been developed for tackling the nearest neighbor search problem:

\begin{enumerate}
    \item \textbf{Space-partitioning trees}: These methods divide the vector space into hierarchical regions, enabling logarithmic search complexity in low dimensions. Notable examples include KD-trees \cite{Bentley1990k}, R-trees \cite{guttman1984r}, VP-trees \cite{yianilos1993}, M-trees \cite{ciaccia1997m}, and Cover trees \cite{beygelzimer2006cover}. A comprehensive overview of these approaches for metric spaces is provided by Zezula et al. \cite{zezula2006similarity}.

    \item \textbf{Mapping-based methods}: These techniques transform the original data into representations that allow for efficient search. This category includes Locality-Sensitive Hashing (LSH) \cite{Indyk1998,Gionis1999,Datar2004,Andoni2006,Andoni2015} and Product Quantization approaches \cite{Jegou2011,Ge2013,Norouzi2013}, which encode vectors into compact codes that approximate distances between points.

    \item \textbf{Navigable graph-based methods}: The most recent and currently state-of-the-art approach that constructs graph structures where vertices represent data points and edges connect similar points. These graphs are designed to allow ``greedy-like'' algorithms to perform directed traversal through the data space.
\end{enumerate}

\subsection{Hierarchical Navigable Small World (HNSW)}

HNSW \cite{hnsw} extends the navigable small-world concept by organizing points into a hierarchy of navigable graphs also named as layers. Each layer contains a progressively smaller subset of the data points. The bottommost layer contains all points, while higher layers form a sparse representation of the data space. 
Graphs of all levels glued together form a navigable small world graph like NSW \cite{nsw2011}. An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.

HNSW supports two primary operations: \textbf{search} (see section \ref{subasec: hnsw-search}) and  \textbf{insertion}. In the Fig. ~\ref{fig:hnsw} is a schematic example of multilayered structure of HNSW.

\begin{figure}
  \centering
  \includegraphics[width=0.5\linewidth]{figs/hnsw-gpt.png}
  \caption{Schematic example HNSW structure. Each layer is a navigable graph that suit for ``greedy-like'' traversal algorithms. Graphs of all levels glued together form a navigable small world graph like similar to NSW. An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.}
\label{fig:hnsw}
\end{figure}

\subsection{Graph Merging as an Iterative Process}
One of the key challenges in maintaining these graph-based indices is efficiently combining data from multiple sources. While HNSW supports dynamic insertions, merging entire graphs as coherent structures presents a different challenge. The quality of the merged graph directly impacts search performance, and naive approaches may lead to suboptimal neighborhood connections or excessive computational costs during the merge operation.

It is worth noting that graph merging can be used as an analog to compaction operations in database systems. In graph-based indices, deletion operations are often implemented by simply marking vertices as deleted without actually removing them, to avoid disrupting the graph's connectivity. This approach prevents the computationally expensive task of rebuilding neighborhoods for affected vertices. Moreover, a true deletion might trigger cascading neighborhood reconstructions that could potentially affect the entire graph, which is unacceptable from a performance perspective. Graph merging provides a way to periodically compact the structure by creating a fresh graph that excludes deleted vertices.

The merge operation can be conceptualized as an iterative process consisting of four key steps:

\begin{enumerate}
    \item \textbf{Processing Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph, and determining which original graph it belongs to.

    \item \textbf{Candidate Collection}: Gathering potential neighbors from both input graphs. This involves searching within each graph structure to find vertices that should be considered for inclusion in $v^*$'s neighborhood.

    \item \textbf{Neighborhood Construction}: Applying a specified strategy (e.g., k-nearest neighbors or relative neighborhood graph) to select the final set of neighbors for $v^*$ from the candidates.

    \item \textbf{Information Propagation}: Deciding what information from the current iteration should be preserved for the next iteration to optimize the process, then proceeding to step 1 with a new vertex.
\end{enumerate}

By carefully designing each of these steps, we can create merge algorithms that balance computational efficiency with the quality of the resulting graph structure.

In this work, we propose three novel algorithms based on this iterative framework—\textbf{Naive Graph Merge (NGM)}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging navigable graphs. These algorithms differ primarily in how they implement each step of the iterative process. For instance, NGM and IGTM select the next vertex for neighborhood construction only from the same graph as the current vertex, promoting locality within original graphs. In contrast, CGTM implements the idea that the next vertex for neighborhood construction can be selected from any input graph, enabling more thorough exploration of the combined space. While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}. These algorithms preserve graph navigability and maintain the structural integrity necessary for fast ANN search while offering different trade-offs between merge accuracy and computational efficiency.

The rest of this paper is organized as follows: Section 2 describes the search algorithms for navigable graphs that we use in merge procedures. Section 3 presents neighborhood construction strategies that are essential for both graph building and merging. Section 4 introduces our three graph merging algorithms in detail. Section 5 presents experimental results comparing the proposed methods. Finally, Section 6 concludes the paper and discusses directions for future research.

\section{Search}

The search process in navigable graphs is a critical operation that underlies both query processing and the graph construction phase. Below, we describe two key search algorithms: \textsc{LocalSearch} for exploring a single graph layer, and \textsc{HNSW-Search} for hierarchical multi-layer traversal.

%

\subsection{Local Search}

\begin{algorithm}
\caption{\textsc{LocalSearch}($G, q, C, k, L$)}\label{alg:local_search}
\textbf{Input:} Graph $G = (V, E)$, query $q \in \mathbb{R}^d$, initial candidate set $C \subset V$, $k \in \mathbb{N}$, $L \in \mathbb{N}$ \\
\textbf{Output:} Approximate $k$-nearest neighbors $V^* \subset V$
\begin{algorithmic}[1]
% \State $C \gets \textsc{InitializeCandidates}()$
\While{True}
    \State $u \gets$ nearest unvisited point to $q$ in $C$
    \State $U \gets \{v \mid (u, v) \in E\}$
    \For{$v \in U$}
        \If{$v$ is unvisited}
            \State $C \gets C \cup \{v\}$
        \EndIf
    \EndFor
    \If{$|C| > L$}
        \State $C \gets$ top $L$ nearest points to $q$ in $C$
    \EndIf
    \If{no updates to $C$}
        \State \textbf{break}
    \EndIf
\EndWhile
\State \Return top-$k$ nearest points to $q$ in $C$
\end{algorithmic}
\end{algorithm}

The \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}) implements a ``greedy'' like search within a single graph layer. Starting with an initial candidate set $C$, it iteratively explores the neighborhood of the closest unvisited point to the query $q$. This exploration strategy balances depth-first search for quick convergence toward the target region with breadth-first search for escaping local minima. In the literature, this algorithm is also named "beam search" \cite{prokhorenkova2020graph}, \cite{yang2024revisiting}. It was introduced in \cite{nsw2011} paper, and later adopted in HNSW in other graph-based methods as a basic search algorithm \cite{diskann}, \cite{wang2021comprehensive}.

For each iteration, \textsc{LocalSearch} selects the closest unvisited point $u$ to the query and explores its immediate neighbors. These neighbors are added to the candidate set if they have not been visited previously. The candidate set is constrained to size $L$ (the expansion factor) by retaining only the $L$ points closest to the query. This pruning step is crucial for an efficient search by focusing computational resources on the most promising candidates.

The search terminates when no additional updates to the candidate set occur during an iteration, indicating convergence. The algorithm then returns the $k$ nearest points to the query from the final candidate set, which represents the approximate $k$-nearest neighbors of the query within the graph structure.

The parameter $L$ provides a direct trade-off between search quality and computational cost. Larger values of $L$ allow for a broader exploration, potentially escaping local minima and finding better global solutions, but at the expense of increased computation time. In the HNSW paper, the parameter $L$ is named as \textbf{ef} (expansion factor) parameter.

\subsection{Hierarchical Search}
\label{subasec: hnsw-search}

\begin{algorithm}
\caption{\textsc{HNSW-Search}($\mathcal{H}, q, v_0, k, L, \ell$)}\label{alg:hnsw_search}
\textbf{Input:} HNSW graph $\mathcal{H} = (G_i)_{i=0}^{l_{\max}}$, query $q \in \mathbb{R}^d$, starting vertex $v_0 \in V$, $k, L \in \mathbb{N}$, search layer $\ell$ \\
\textbf{Output:} Approximate $k$-nearest neighbors $V^* \subset V$
\begin{algorithmic}[1]
\State $v^* \gets v_0$ 
\For{$i = l_{\max} \textbf{ down to } \ell$}
    \State $v^* \gets \textsc{LocalSearch}(G=G_i, q=q, C=\{v^*\}, k=1, L=L)$
\EndFor
\State \Return \textsc{LocalSearch}$(G_\ell, q, \{v^*\}, k, L)$
\end{algorithmic}
\end{algorithm}

The \textsc{HNSW-Search} algorithm (Algorithm~\ref{alg:hnsw_search}) leverages the hierarchical structure of HNSW to efficiently navigate through the vector space. The search begins at the highest layer $l_{\max}$ of the HNSW structure with a single entry point $v_0$. At each layer, \textsc{LocalSearch} is used to find the best approximation of the nearest neighbor to the query within that layer.

The algorithm traverses from the highest layer down to the target layer $\ell$, using the result from each layer as the entry point for the next lower layer. This coarse-to-fine approach allows the search to quickly focus on the relevant region of the vector space at higher, sparser layers before refining the search in the more densely connected lower layers.

At each layer $i$ (from $l_{\max}$ down to $\ell+1$), \textsc{LocalSearch} is executed with $k=1$ to find a single nearest neighbor to the query. This single neighbor serves as an entry point for the subsequent layer. At the target layer $\ell$, a final \textsc{LocalSearch} is performed with the specified value $k$ to retrieve the $k$-nearest neighbors.

This hierarchical search significantly reduces the number of distance computations compared to a flat graph search, especially for large-scale datasets. The efficiency comes from the progressive narrowing of the search space as the algorithm moves down through the layers, focusing the search on increasingly relevant regions.

Both search algorithms are fundamental not only for answering queries but also for the merge operations described in subsequent sections, as they provide the mechanism for identifying candidate neighbors when reconstructing connections in the merged graph.

\textsc{HNSW-Search} algorithm is a default search algorithm that is exposed to the user in all popular HNSW implementations with a default value of the parameter $\ell = 0$, and assuming that $v_0$ is one of the vertex of the top layer \cite{Vardanian_USearch_2023}, \cite{hnswlib2025}, \cite{faiss}.

In the text, we also refer to the \textsc{HNSW-Search} algorithm as a ``standard'', or ``default search" algorithm. The \textsc{LocalSearch} function is hidden from the user and is used only as a part of a particular implementation".

\section{Neighborhood Construction Strategies}

\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}($v^*, C, k$)}\label{alg:knntrategy}
\textbf{Input:} Vertex $v^*$, candidate set $C$, number of neighbors $k$ \\
\textbf{Output:} Set $C'$ of $k$ closest neighbors
\begin{algorithmic}[1]
\State $C' \gets$ $k$-nearest neighbors of $v^*$ in $C$
\State \Return $C'$
\end{algorithmic}
\end{algorithm}

Various strategies exist for selecting a node's neighborhood to form a navigable property for a graph. Our merging algorithms are designed to be agnostic to the specific neighborhood construction method, which can be provided as a parameter. We present two representative strategies below.

The \textsc{KNN-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:knntrategy}) implements the simplest approach, where a vertex's neighborhood consists of its $k$-nearest neighbors from the candidate set. This strategy prioritizes proximity, but may not optimize for graph navigability properties.

In contrast, the \textsc{RNG-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:rngstrategy}) implements a more sophisticated approach based on relative neighborhood graph principles. It processes candidates in ascending order of distance to the target vertex $v^*$. For each candidate $v$, it checks whether $v$ is closer to $v^*$ than to any previously selected neighbor $w$. This pruning condition helps create better-connected graphs with improved navigability by ensuring that edges span different directions in the vector space rather than clustering in the same region. The algorithm limits the neighborhood size to at most $m$ vertices to control maximum vertex degree.

The choice of neighborhood construction strategy significantly impacts both search performance and the computational cost of index maintenance operations, including merges. While \textsc{KNN-Neighborhood-Construction} is computationally simpler, \textsc{RNG-Neighborhood-Construction} typically produces graphs with better search performance at the expense of more complex neighborhood formation. In this paper, we performed our experiments with \textsc{RNG-Neighborhood-Construction}.

\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}($v^*, C, m$)}\label{alg:rngstrategy}
\textbf{Input:} Vertex $v^*$, candidate set $C$, maximum neighborhood size $m$ \\
\textbf{Output:} Filtered neighbor set $C'$
\begin{algorithmic}[1]
\State Sort $C$ in ascending order by distance to $v^*$
\State $C' \gets \emptyset$
\For{$v \in C$}
    \State $f \gets$ true
    \For{$w \in C'$}
        \If{$\rho(v^*, v) \geq \rho(v, w)$}
            \State $f \gets$ false
            \State \textbf{break}
        \EndIf
    \EndFor
    \If{$f$}
        \State $C' \gets C' \cup \{v\}$
    \EndIf
    \If{$|C'| \geq m$}
        \State \textbf{break}
    \EndIf
\EndFor
\State \Return $C'$
\end{algorithmic}
\end{algorithm}

% \textsc{KNN-Neighborhood-Construction}(Algorithm~\ref{alg:knntrategy}) is a simplest way to form neighbors, while \textsc{RNG-Neighborhood-Construction} (Algorithm~\ref{alg:rngstrategy}) is more advanced, that paratactically provides better navigability property.  
\section{Merge Algorithms}

%We provide 3 merge algorithms: "Naive Merge", "IGTM", and "CGTM". We also refer to these algorithm as layer-merge algorithms. "Naive Merge" is a straight forward method, while "IGTM" and "CGTM" are more advanced. Together, these two algorithms constitute the main contribution of the work. The HNSW structure is a sequence of graphs, called layers. We start from description of a general algorithm for merging multi layers structure.
We now present three algorithms for merging HNSW graphs: \textsc{SIGM},  \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}. These algorithms operate layer by layer. We first describe the general procedure for merging the multi-layer HNSW structure, followed by the specific logic for merging a single layer (\textit{layer-merge algorithms}).

\subsection{Simple Insertion Graph Merge}
Before we start, we need provide a short descriptopn of simple insertion graph merge (SIGM). It is a default merging strategy that uses an insertion operation to merge graphs. It takes the largest graph and sequentially inserts data from other graph. Actually, it is not merging procedure, but we use it as a basic benchmark.

\subsection{HNSW General Merge Framework}

The \textsc{HNSW-General-Merge} algorithm (Algorithm~\ref{alg:general_merge}) provides a framework for merging two HNSW structures, $\mathcal{H}_a$ and $\mathcal{H}_b$. It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{NGM}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$. The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform searches using \textsc{HNSW-Search}.

\begin{algorithm}
\caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b, \text{LayerMergeAlgo}, \text{params}$)}\label{alg:general_merge}
\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b $. Chosen layer merge algorithm $\text{LayerMergeAlgo}$ (e.g., \textsc{NGM}). Algorithm-specific parameters $\text{params}$. \\
\textbf{Output:} Merged HNSW graph $\mathcal{H}_c = (G^c_i)_{i=0}^{l_{\max}^c}$
\begin{algorithmic}[1]
\State $l_{\max}^a \gets \mathcal{H}_a.\text{getMaxLayerNumber()}$
\State $l_{\max}^b \gets \mathcal{H}_b.\text{getMaxLayerNumber()}$
\State $l_{\max}^c \gets \max(l_{\max}^a, l_{\max}^b)$
\For{$i = 0 \textbf{ to } l_{\max}^c$} \Comment{Merge each layer}    
    \State $G^c_i \gets \text{LayerMergeAlgo}(\mathcal{H}_a, \mathcal{H}_b, \ell=i, \text{params})$ \Comment{Calls e.g., Merge-Naive}
\EndFor
\State $\mathcal{H}_c \gets (G^c_i)_{i=0}^{l_{\max}^c}$ % Construction is implicit in the loop

\State \Return $\mathcal{H}_c$
\end{algorithmic}
\end{algorithm}

Note that the merging of different layers (iterations of the loop in Algorithm~\ref{alg:general_merge}) is independent and can potentially be parallelized.

%\subsection{Naive Merge}
%The Naive Merge algorithm \ref{alg:merge_naive} surprisingly is very simple. It takes one vertex $v^*$ by one from the graph $G^a$. Then algorithm prepares the candidate set $\mathcal{C}$ for a new neighborhood (lines 6,7). It uses \textsc{HNSW-Search} to get $m$-closest vertices to $v^*$ in the graph $G^b$, and combines with an old neighborhood of the vertex $v^*$. After that in the line 8 the candidate set $\mathcal{C}$ is passed into the neighborhood construction function, which selects the vertices for the new merged neighborhood for $v^*$.
%Lines 9-12 contains the same logic for the forming new neighborhoods of graph $G^b$

\subsection{Naive Graph Merge}
The \textsc{Naive Graph Merge} (\textsc{NGM}) algorithm (Algorithm~\ref{alg:merge_naive}) provides a straightforward method for merging a single layer $\ell$. The algorithm begins by extracting the target layers from both input HNSW structures and initializing the merged graph's vertex set as the union of vertices from both input graphs (lines 1-4).

For each vertex $v^*$ in graph $G^a_\ell$ (line 5), the algorithm:
\begin{enumerate}
  \item Searches for potential neighbors in graph $G^b_\ell$ using \textsc{HNSW-Search} (line 6)
  \item Combines these candidates with $v^*$'s original neighbors from $G^a_\ell$ (line 7)
  \item Selects the final neighborhood for $v^*$ using the specified \textsc{NeighborhoodConstruction} strategy and adds the corresponding edges to $E^c$ (line 8)
\end{enumerate}

The same process is then repeated for all vertices of graph $G^b_\ell$ (lines 9-12).

This approach ensures that each vertex in the merged graph has an appropriate neighborhood that incorporates information from both input graphs. However, it is computationally intensive due to repeated \textsc{HNSW-Search} calls that traverse multiple layers for each vertex.

\begin{algorithm}
\caption{\textsc{NGM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{NeighborhoodConstruction}, m, \text{search\_ef}, v_{entry}$)}\label{alg:merge_naive}
\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b$; target layer $\ell$; Neighborhood construction function \textsc{NeighborhoodConstruction}; target neighborhood size $m$; search parameter $\text{search\_ef}$; entry point $v_{entry}$ (e.g., from $\mathcal{H}_a$ or $\mathcal{H}_b$) \\
\textbf{Output:} Merged graph $G^c$
\begin{algorithmic}[1]
\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
% \State $m \gets \dots$ % Now passed as input parameter

\For{$v^* \in V^a$} 
    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_b, q=v^*, v_{entry}=v_{entry}^b, k=m, L=\text{search\_ef}, \ell=\ell)$ 
    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^a \} \cup \mathcal{C}^b$  
    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\For{$v^* \in V^b$} 
    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_a, q=v^*, v_{entry}=v_{entry}^a, k=m, L=\text{search\_ef}, \ell_{target}=\ell)$ 
    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^b \} \cup \mathcal{C}^a$  
    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\State \Return $G^c = (V^c, E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Intra Graph Traversal Merge}
The most effort of the \textsc{NGM} algorithm lies in obtaining the set of neighborhood candidates from the other graph utilising \textsc{HNSW-Search} procedure, that every time traverses layer graphs from the top-level down to layer number $\ell$. The number of computations can be reduced if we select the next vertex to process $v^*$ close to the previous one (line 15), instead of randomly choosing it. Thus, for the new $v^*$ the neighborhood candidates will also be close to the previous candidates set. To search these new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traverses the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (lines 11). In line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
In lines 15,16 we select new processing vertex $v^*$  close to the previous processing vertex $v^*$ that was not already processed.  
To ensure that the new processing vertex $v^*$ is not very far from the previous $v^*$ we bound the size of the results of the \textsc{LocalSearch} procedure controlling by next\_step\_k parameter. Once \textsc{LocalSearch} can't find enough close not processed vertex (if condition in line 17), in line 7 we choice new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly. \\
After we have processed all vertices from the graph $G^a$. We do the same for the vertices of graph $G^b$ (lines 21-22).

\begin{algorithm}
\caption{\textsc{IGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:IGTM}
\textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the merging layer number $\ell$, the size of the forming neighborhoods $m \in \mathbb{N}$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
\textbf{Output:}  Merged graph $G^c$ 
\begin{algorithmic}[1]
\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
\State $\mathcal{V}_{not\_done} \gets V^a$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{random choice from } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^b  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k=M, L=\text{jump\_ef}, \ell)$

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $\mathcal{C}^b  \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$

        \State $E^c \gets E^c \cup  \{ (v^*, v)  : v \in \textsc{NeighborhoodConstruction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{P}^b \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_M \} $

        \State $\mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\{v^*\} , k=\text{next\_step\_k}, L=\text{next\_step\_ef})$

        \State $\mathcal{C}^a \gets \mathcal{C}^a \cap \mathcal{V}_{not\_done}$

        \If{$\mathcal{C}^a = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \mathcal{C}^a_1$
    \EndWhile
\EndWhile
\State $\mathcal{V}_{not\_done} \gets V^b$
\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State Repeat the same process for $V^b$ with the roles of $\mathcal{H}_a$ and $\mathcal{H}_b$ swapped.
\EndWhile
\State \Return $G^c=(V^c,E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Cross Graph Traversal Merge}

Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilizes \textsc{LocalSearch} procedure to reduce computation efforts. The difference is that \textsc{IGTM} algorithm chooses the next processing vertex $v^*$ from the same graph, while \textsc{CGTM} looks for the new processing vertex $v^*$ in both graphs $G^a$, and $G^b$. 
Thus in line 24 $v^*$ is chosen from the set $\mathcal{C}_{not\_done}$, which is one of the not processed vertex of both graph (line 21). The intuition underlying \textsc{CGTM} algorithm is that when we choose a new processing vertex $v^*$ from both graphs, we reduce the number of times when $v^*$ is chosen randomly, thus minimizing the number of times that we use the more expensive search procedure \textsc{HNSW-Search}.

\begin{algorithm}
    \caption{\textsc{CGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:CGTM}
    \textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, the merging layer $\ell$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, neighborhood sizes $M, m \in \mathbb{N}$ \\
    \textbf{Output:} Merged graph $G^c$ 
\begin{algorithmic}[1]

\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$
% \State $m \gets \mathcal{H}_a.m_0 \text{ if } \ell = 0 \text{ else } \mathcal{H}_a.m$
\State $\mathcal{V}_{not\_done} \gets V^a \cup V^b$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{randomly choice from } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^a  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^a, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \State $\mathcal{P}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $ \mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\mathcal{P}^a  , k=m, L=\text{local\_ef})$

        \State $\mathcal{C}^b \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \If{$v^* \in V^a $}
            \State $\mathcal{C}\gets  \{v : (v^*, v) \in E^a \} \cup  \mathcal{C}^b\}$
        \Else
            \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^b \} \cup  \mathcal{C}^a\}$
        \EndIf

        \State $E^c \gets E^c \cup \{ (v^*, v) : v \in \text{neighborhood\_construction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{C}^a_{not\_done} \gets \{\mathcal{C}^a_1, \mathcal{C}^a_2, ..., \mathcal{C}^a_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}^b_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \}  \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}_{not\_done} \gets \mathcal{C}^a_{not\_done} \cup \mathcal{C}^b_{not\_done}$

        \If{$\mathcal{C}_{not\_done} = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$

        \State $\mathcal{P}_a \gets \mathcal{C}^a$
        \State $\mathcal{P}_b \gets \mathcal{C}^b$
    \EndWhile
\EndWhile
\State \Return $G^c=(V^c,E^c)$
\end{algorithmic}
\end{algorithm}

\section{Computational Experiments}

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
  \caption{Recall vs distance count on search stage for merged graphs. }
\label{fig:search}
\end{figure}

% We implemented proposed algorithms using python. For experiments we use Sift1m dataset divided into two subset of 500k vectors at each. Our algorithm merge that two subsets of 500k vectors. In order to be independent of the implementation details and low level optimisation, we do comparison of the algorithms based on the number of distance computation during the merge process. \ref{fig:sfig1} shows that \textsc{IGTM} about 5 times less computations than naive algorithm. Surprisingly \textsc{CGTM} algorithm do slightly more computations than \textsc{IGTM}, nevertheless it significantly outperform \textsc{NGM} algorithm.\\

% The other important characteristics of merging process is how accurate the merge is done. It can be easily verified for the case of \textsc{KNN-Neighborhood-Construction} procedure, by comparing the set of merged neighborhoods with k-nearest-neighbor set for each vertex. However to follow the same way in the case of \textsc{RNG-Neighborhood-Construction} is not good idea, because the \textsc{RNG-Neighborhood-Construction} is one of the possible heuristic for Voronoi-cell approximation. Therefore we measure the recall-distance computation trade off of \textsc{HNSW-Search} procedure on the HNSW graphs before merging, and after.

% As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{IGTM}, and  \textsc{CGTM} has.

\subsection{Setup}

We evaluated the proposed merge algorithms on the standard ANN benchmark dataset SIFT1M of 1 million 128-dimensional vectors. We divided it into two disjoint subsets of 500k vectors. On those subsets we built two HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ using the following parameters: $M = 16$, $M0$ = 32, ef\_construction = 32.
These indices are then merged using simple insertion strategy \textsc{SIGM} and the three proposed algorithms: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}. We keep the same values of the parameters $m = 16$, and $m0 = 32$ as the target neighborhood size in the merged graph.

In order to be independent of implementation details and low-level optimisations, we made a comparison of the algorithms based on the number of distance computations required for the merge process.

Other important characteristics of the merge process is how accurate the merge is done. In this case, the word ``merging accuracy'' can mean different things. For example, we can estimate merging accuracy as how similar the neighborhoods of the merged graph to the neighborhoods of the graph constructed by simple insertion strategy \textsc{SIGM}. 
Another way, which we prefer, is to verify how the merged graph is good for search. To quantify search quality, we performed a standard search performance test on the merged graphs. So, for the merged graphs, we have measured a trade-off between recall and the number of distance computations, by running algorithm ~\ref{alg:hnsw_search} with different values of the parameter $L$ (search expansion factor). We measured the recall@5 metric for L = 32, 40, 50, 64, and 72. This metric indicates how often the true nearest neighbors appear in the top-5 results returned by the search algorithm.

Formally, for a given query point \(q\), let \(P_k(q)\) denote the set of \(k\) true nearest neighbors, and let \(A_k(q)\) be the set of top \(k\) points returned by the algorithm. Then, the recall@k is defined as:

\[
\text{recall@}k = \frac{|P_k(q) \cap A_k(q)|}{k}
\]

In our experiment as recall@5 we report an averaged value over all sequences of searches.

\subsection{Results}

The searching quality of the merged graph represented in Fig. ~\ref{fig:search}. The graph plots trade-off between recall and the number of distance computations. As can be seen, \textsc{SIGM} with ef\_construction $\ge$ 32 and \textsc{NGM} with parameter jump\_ef = 20 produce merged graph with slightly better searing quality than other algorithms.

Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the search algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $L$ are almost the same.  Therefore, the graph is better if it provides better search recall for a fixed value of the parameter $L$. Taking this into account, we can better interpret Fig. ~\ref{fig:recall_vs_merge}. It is easy to see the following:

\begin{itemize}
    \item\textsc{IGTM} and \textsc{IGTM}  in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24;
    \item \textsc{NGM} and \textsc{SIGM} perform the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood;
    \item \textsc{CGTM} achieving comparable recall with approximately 60\% fewer computations than \textsc{NGM} and \textsc{SIGM};
    \item \textsc{IGTM} provides the best run-time performance, reducing distance computations about 20\% than \textsc{CGTM}, and about 70\% than \textsc{NGM} and \textsc{SIGM} while achieving comparable recall.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{figs/recall_vs_merge_distance_count.pdf}
  \caption{Recall of final merged graph vs merging efforts}
  \label{fig:recall_vs_merge}
\end{figure}

\section{Conclusion and Future Work}

In this work, we introduced three algorithms to merge hierarchical navigable small-world graphs: \textsc{NGM} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).

The \textsc{NGM} algorithm provides a straightforward but computationally intensive approach, performing standard HNSW searches to find sets of candidates to reconstruct the neighborhood of a vertex chosen in an arbitrary way. \textsc{IGTM} and \textsc{CGTM} improve efficiency by leveraging locality—processing vertices close to each other sequentially and using the less expensive \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}).

Our experimental results demonstrate that both \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy. Our evaluation on the dataset of 1 million 128-dimensional vectors (SIFT1M) shows that \textsc{IGTM} and \textsc{CGTM} are more than 3 times faster than the straightforward insertion strategy \textsc{SIGM}, with minimal impact on recall performance.

Interestingly, our experiments revealed that \textsc{IGTM} outperformed \textsc{CGTM} in terms of computational efficiency, contrary to our initial expectations. We had anticipated that \textsc{CGTM} would be more efficient since it can select the next vertex for neighborhood construction from both graphs, theoretically reducing the chances of getting stuck and thus requiring fewer costly standard search operations. However, it appears that the overhead of selecting the next processing vertex, for which a neighborhood is forming, in \textsc{CGTM} is too computationally expensive, and these costs are not offset by the reduction in searches when ``getting stuck'' with selecting the next close vertex to process. The study of this fact is a subject for further work.

In addition, an important direction for future work is adapting the proposed merge algorithms to handle deleted vertices. This would enable their use in compaction processes, where graphs are periodically restructured to remove obsolete entries and maintain search efficiency. By extending our merge algorithms to filter out deleted nodes during the ``Processing Vertex Selection'' phase.

Also, the researchers can focus on the idea that the neighborhood construction can be done for the set of vertex close to each other instead of forming neighborhood for one vertex per iteration. This can be more suitable for GPU or NPU settings.

Additional areas for exploration include adaptive parameter selection based on dataset characteristics, and extensions to other graph-based index structures beyond HNSW.

Finally, future work can be applied to develop a fast graph construction procedure using one of the merge algorithms. It seems very natural to start from the set of small graphs and merge them in a recursive manner.

\bibliographystyle{plain}
\bibliography{references}
\end{document}